---
title: "Analysis"
format: html
editor: visual
author: 'Yifan Wu, Hongyu Jing'
---

## Input dataset

```{r}
library(readxl)
library(tidyverse)
dat<- read_xls("SAP2020_merged_v1.xls")
<<<<<<< HEAD
dat <- as.data.frame(dat)
for (i in c(1:7)) {
  dat[,i]=factor(dat[,i])
}

dat
```

-   We can visualize the potential spatial effect on plant growth, indicated by plant height or flowering time, by plotting the field row and column layout and color-coding with the phenotypes.

```{r}
ggplot(data=dat, aes(x=Row, y=Column, color=EstimatedPlotYield))+
  geom_point()+
  scale_colour_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 900,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "colour")

```

-   We can investigate the relationship between plant phenotypes, for example plotting 2d histograms and scatterplots bewtween each two agronomic traits.

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y, use = "complete")
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste(prefix, txt, sep="")
  if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex * abs(r))
}

dat_num <- dat[,unlist(lapply(dat, is.numeric))]  
yield_traits <- c("PaniclesPerPlot", "PanicleGrainWeight", "EstimatedPlotYield")
dat_yield <- dat[,yield_traits]
pairs(dat_yield,  upper.panel=panel.cor)
```
=======
dat[1:5,]
```

The dataset is the agronomic traits that were measured in a sorghum field experiment with two levels of nitrogen treatment and around 400 genotypes. Hopefully we will also find the sub-population information of these genotypes as an additional categorical attribute.\

Each row represents a small pieces of land with it's unique row number and column number which contains around 10 individual sorghum plants. The explanation for each column stated below:\

**Plot ID&Sorghum Accession:** Unique ID for each plants\

**SNPDataID：** Unique ID\

**Row&Column:** spatial data for sorghum plants\

**Block:** Blocking number\

**Treatment:** two levels of nitrogen treatment\

**DaysToBloom:** How many days until the sorghum flower bloom(unit:day)\

**MedianLeafAngle:** Median value of leaf angle for each 10 individual sorghum plants(unit:degree)\

**LeafAngleSDV:** Standard deviation of leaf angle for 10 individual sorghum plants\

**PoorStand:** If half of the sorghum plants are poor stand, if yes then the answer is Y\

**PaniclesPerPlot:** How many panicles per plot have\

**PanicleGrainWeight:** How much grain weight for each panicle(unit:gram)\

**EstimatedPlotYield:** The estimated plot yield for 10 individual sorghum plants(unit:gram)\

**FlagLeafLength:** The mean value of the flag leaf's length for each 10 individual sorghum plants(unit: cm)\

**FlagLeafWidth:** The mean value of the flag leaf's width for each 10 individual sorghum plants(unit: cm)\

**ExtantLeafNumber:** The number of visible leaves on the main stalk at maturity stage.\

**PlantHeight:**The mean value of the plant height for each 10 individual sorghum plants(unit: cm)\

**ThirdLeafLength:** The mean value of the third leaf's length for each 10 individual sorghum plants(unit: cm)\

**ThirdLeafWidth:** The mean value of the third leaf's width for each 10 individual sorghum plants(unit: cm)\

## data cleaning

```{r}
##delete the useless column

newdat<- dat[,c(-1,-3)]

##there exist missing data in several rows at different column, we will rearrange it during analysis
```

## data exploring

```{r}
summary(newdat)
```
## Data analysis
### 1. check the distribution for numeric output column
#### 1.1 box plot for leaf dimension
```{r}
library(ggplot2)
library(dplyr)
require(reshape2)
library(tidyr)


###select leaf dimention data to plot a box plot
###select required column
analysis_le<- newdat[,c(13,14,17,18)]
###clean navalue
clean_le<- analysis_le%>%
  melt()%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_le,aes(x=variable, y=value))+
  geom_boxplot()+
  labs(
    title = "box plot for leaf dimension"
  )


  

```
Based on the plot above, we can see that overall the dimension of the third leaf is larger than the flag leaf which is reasonable.
#### 1.2 histogram for blooming day

```{r}
##create some histogram for each variable to have a basic understand for each variable

###blooming date
analysis_bloom<-melt(newdat[,c(6)])

###clean navalue
clean_bloom<- analysis_bloom%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_bloom,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "histogram for bloomingday"
  )


```
Based on the histogram above, we know that usually it takes around 50 to 80 days for the sorghum flower bloom.The highest probability range is around 60 to 70 days.
#### 1.3 histogram for median leaf angle
```{r}

analysis_angle<-melt(newdat[,c(7)])

###clean navalue
clean_angle<- analysis_angle%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_angle,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "histogram for median leaf angle"
  )

```
Based on the histogram above, we know that usually the median value of leaf angle for each 10 individual sorghum plants is around 40 to 45 degree.
#### 1.4 histogram for panicle amount
```{r}

analysis_panicle<-melt(newdat[,c(10)])

###clean navalue
clean_panicle<- analysis_panicle%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_panicle,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "histogram for panicle amount"
  )

```
Based on the histogram above, we know that usually the panicles per plot have is around 0 to 20 where the highest possibility is around 13.
####1.5 histogram for grain weight
```{r}

analysis_grain<-melt(newdat[,c(11)])

###clean navalue
clean_grain<- analysis_grain%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_grain,aes(x=value))+
  geom_histogram(binwidth = 5)+
  labs(
    title = "histogram for grain weight"
  )

```
Based on the histogram above, we know that usually the grain weight for each panicle is around 0 to 50 grams.
####1.6 histogram for yield
```{r}

analysis_yield<-melt(newdat[,c(12)])

###clean navalue
clean_yield<- analysis_yield%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_yield,aes(x=value))+
  geom_histogram(binwidth = 50)+
  labs(
    title = "histogram for yield"
  )

```

Based on the histogram above, we know that the estimated plot yield for 10 individual sorghum plants is around 0 to 1000 grams where the highest possibility range is around 300 to 500 gram.
####1.7 histogram for plant height
```{r}

analysis_height<-melt(newdat[,c(16)])

###clean navalue
clean_height<- analysis_height%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_height,aes(x=value))+
  geom_histogram(binwidth = 10) +
  labs(
    title = "histogram for plant height"
  )

```
Based on the histogram above, we know range for plot height is around 30 to 300cm where usually the height will be around 70 cm.
###2.Explore the effect of the nitrogen treatment for leaf dimension/plant height/yield
####2.1 plant height
```{r}
effect_height<-newdat[,c(5,16)]
###clean navalue
clean_eff_height<- effect_height%>%
  ###cleaning na value
  drop_na(PlantHeight)

ggplot(clean_eff_height, aes(x=Treatment, y=PlantHeight))+
  geom_boxplot() +
  labs(
    title = "plant height by group"
  )
ggplot(clean_eff_height, aes(x=PlantHeight, fill=Treatment))+
  geom_histogram() +
  labs(
    title = "plant height by group"
  )
```
Based on the boxplot above, there is no significant difference effect of the treatment we can see. The histogram also support this result.


#### 2.2 yield
```{r}
effect_yield<-newdat[,c(5,12)]
###clean navalue
clean_eff_yield<- effect_yield%>%
  ###cleaning na value
  drop_na(EstimatedPlotYield)

 
ggplot(clean_eff_yield, aes(x=Treatment, y=EstimatedPlotYield))+
  geom_boxplot() +
  labs(
    title = "yield by treatment"
  )


```
Based onn the boxplot above, we can see that sufficient nitrogen can increase the estimated plot yield even include the outliers.

####2.3 leaf length
```{r}
###select required row
leaf_length<- newdat[,c(5,13,17)]
leaf_length %>% 
  rename( # here we rename the columns so things look nice in the graph later
    Flag = FlagLeafLength,
    Third = ThirdLeafLength
  ) %>% 
  pivot_longer( # then we collapse the columns for each side of the brain into a single column, with a second column holding size values
    cols = c("Flag", "Third"),
    names_to = "Side",
    values_to = "Size"
  ) %>%
  ###clean na value
  drop_na(Size)%>% # then we plot and give it a title
  ggplot(
    aes(
      x = Side,
      y = Size,
      fill = Treatment
    )
  ) + 
  geom_boxplot() +
  labs(
    title = "leaf length by group"
  )

```
Based on the box plot above, we can only say the sufficient nitrogen may have a little influence on leaf length.

####2.4 leaf width
```{r}
###select required row
leaf_width<- newdat[,c(5,14,18)]
leaf_width %>% 
  rename( # here we rename the columns so things look nice in the graph later
    Flag = FlagLeafWidth,
    Third = ThirdLeafWidth
  ) %>% 
  pivot_longer( # then we collapse the columns for each side of the brain into a single column, with a second column holding size values
    cols = c("Flag", "Third"),
    names_to = "Side",
    values_to = "Size"
  ) %>%
  ###clean na value
  drop_na(Size)%>% # then we plot and give it a title
  ggplot(
    aes(
      x = Side,
      y = Size,
      fill = Treatment
    )
  ) + 
  geom_boxplot() +
  labs(
    title = "leaf width by group"
  )

```
Based on the boxplot above, we can say that the sufficient nitrogen has no influence on flag leaf width and it may has negative influence on third leafwidth.



>>>>>>> f0f4bdd27f4bddf47cbca475cb4898ef69ea37ab
