---
title: "Analysis"
format: html
editor: visual
author: 'Yifan Wu, Hongyu Jin'
output: html_document 
runtime: shiny
---

## Input dataset

```{r}
library(readxl)
library(tidyverse)
dat<- read_xls("SAP2020_merged_v1.xls")

dat[1:5,]
```

The dataset is the agronomic traits that were measured in a sorghum field experiment with two levels of nitrogen treatment and around 400 genotypes. Hopefully we will also find the sub-population information of these genotypes as an additional categorical attribute.\

Each row represents a small pieces of land with it's unique row number and column number which contains around 10 individual sorghum plants. The explanation for each column stated below:

| Column name               | Description                                                                     |     Unit |
|-----------------|:-------------------------------------|----------------:|
| Plot ID&Sorghum Accession | Unique ID for each plants                                                       | Constant |
| SNPDataID                 | Unique ID                                                                       |        / |
| Row&Column                | spatial data for sorghum plants                                                 | Constant |
| Block                     | Blocking number                                                                 | Constant |
| Treatment                 | two levels of nitrogen treatment                                                |        / |
| DaysToBloom               | How many days until the sorghum flower bloom(unit:day)                          |     days |
| MedianLeafAngle           | Median value of leaf angle for each 10 individual sorghum plants(unit:degree)   |   degree |
| LeafAngleSDV              | Standard deviation of leaf angle for 10 individual sorghum plants               | Constant |
| PoorStand                 | If half of the sorghum plants are poor stand, if yes then the answer is Y       |      Y/N |
| PaniclesPerPlot           | How many panicles per plot have                                                 | Constant |
| PanicleGrainWeight        | How much grain weight for each panicle                                          |     Gram |
| EstimatedPlotYield        | The estimated plot yield for 10 individual sorghum plants                       |     Gram |
| FlagLeafLength            | The mean value of the flag leaf's length for each 10 individual sorghum plants  |       cm |
| FlagLeafWidth             | The mean value of the flag leaf's width for each 10 individual sorghum plants   |       cm |
| ExtantLeafNumber          | The number of visible leaves on the main stalk at maturity stage.               | Constant |
| PlantHeight               | The mean value of the plant height for each 10 individual sorghum plants        |       cm |
| ThirdLeafLength           | The mean value of the third leaf's length for each 10 individual sorghum plants |       cm |
| ThirdLeafWidth            | The mean value of the third leaf's width for each 10 individual sorghum plants  |       cm |

## Data cleaning

```{r}
##delete the useless column

newdat<- dat[,c(-1,-3)]

##there exist missing data in several rows at different column, we will rearrange it during analysis
```

## Data exploring

```{r}
summary(newdat)
```

## Data analysis

### 1. check the distribution for numeric output column

#### 1.1 box plot for leaf dimension

```{r}
library(ggplot2)
library(dplyr)
require(reshape2)
library(tidyr)


###select leaf dimention data to plot a box plot
###select required column
analysis_le<- newdat[,c(13,14,17,18)]
###clean navalue
clean_le<- analysis_le%>%
  melt()%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_le,aes(x=variable, y=value))+
  geom_boxplot()+
  labs(
    title = "Box plot for flag and third leaf dimension",
    x='Dimension variables',
    y='Leaf dimension(Unit: cm)'
  ) 


  

```

The first thing we want to explore is if there exist some difference between the leaf dimension in flag leaf and third leaf. Thus, we got a boxplot for flag and third leaf.

Based on the plot above, we can see that overall the dimension of the third leaf is larger than the flag leaf which is reasonable. Also, we noticed that the outliers for flag leafs are all occur in the top for both length and width, same for the third leaf which the outliers occur in the both side. Thus we thought if the leaf length has the relationship with leaf width.

```{r}
##clean NA value
corr_ana<- na.omit(analysis_le)

##plot a scatter plot first to check
ggplot(corr_ana,aes(x=FlagLeafLength, y=FlagLeafWidth))+
  geom_point()+
  labs(
    title = "Scatter plot for flag leaf dimension",
    x='Flag leaf Length (Unit: cm)',
    y='Flag leaf width (Unit: cm)'
  )
ggplot(corr_ana,aes(x=ThirdLeafLength, y=ThirdLeafWidth))+
  geom_point()+
  labs(
    title = "Scatter plot for third leaf dimension",
    x='Third leaf Length (Unit: cm)',
    y='Third leaf width (Unit: cm)'
  )
  
 ###combine these two plot together
Leaf_type <- c("Third leaf" = "#619CFF", "Flag leaf" = "#F8766D")
  ggplot(corr_ana)+
  geom_point(aes(x=ThirdLeafLength, y=ThirdLeafWidth, colour='Third leaf'))+
  geom_point(aes(x=FlagLeafLength, y=FlagLeafWidth, colour='Flag leaf'))+
  labs(
    title = "Scatter plot for flag and third leaf's dimension",
    x='leaf Length (Unit: cm)',
    y='leaf width (Unit: cm)'
  )+
     scale_color_manual(values = Leaf_type)
```

Based on the two scatter plot above, we can briefly check if there is relationship between leaf length and leaf width. For the flag leaf dimension plot, we can easily find a ellipse which shows there is positive relationship between flag leaf length and flag leaf width. For the third leaf dimension plot, the ellipse is hard to find and it more like a circle which means the relationship between third leaf length and width may not so large. Thus, further analysis needs to be done to draw a clear conclusion.

#### 1.2 histogram for blooming day

```{r}
##create some histogram for each variable to have a basic understand for each variable

###blooming date
analysis_bloom<-melt(newdat[,c(6)])

###clean navalue
clean_bloom<- analysis_bloom%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_bloom,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "Histogram for blooming day",
       x='Blooming time(Unit: day)',
    y='Counts of lands'
  )


```

Then, we want to check the distribution for multiple variables to have a basic understanding for each variaible. Thus, we create a histogram for each one and check it's distribution.

Based on the histogram above, we know that usually it takes around 50 to 80 days for the sorghum flower bloom.The highest probability range is around 60 to 70 days.\

#### 1.3 histogram for median leaf angle

```{r}

analysis_angle<-melt(newdat[,c(7)])

###clean navalue
clean_angle<- analysis_angle%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_angle,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "Histogram for median leaf angle",
       x='Leaf angle(Unit:degree)',
    y='Count of lands'
  )

```

Based on the histogram above, we know that usually the median value of leaf angle for each 10 individual sorghum plants is around 40 to 45 degree.\

#### 1.4 histogram for panicle amount

```{r}

analysis_panicle<-melt(newdat[,c(10)])

###clean navalue
clean_panicle<- analysis_panicle%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_panicle,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "Histogram for panicle amount",
       x='Panicle amount',
    y='Number of lands'
  )

```

Based on the histogram above, we know that usually the panicles per plot have is around 0 to 20 where the highest possibility is around 13.\

#### 1.5 histogram for grain weight

```{r}

analysis_grain<-melt(newdat[,c(11)])

###clean navalue
clean_grain<- analysis_grain%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_grain,aes(x=value))+
  geom_histogram(binwidth = 5)+
  labs(
    title = "Histogram for grain weight",
       x='Average grain Weight(Unit: gram)',
    y='Number of lands'
  )

```

Based on the histogram above, we know that usually the grain weight for each panicle is around 30 to 50 grams.\

#### 1.6 histogram for yield

```{r}

analysis_yield<-melt(newdat[,c(12)])

###clean navalue
clean_yield<- analysis_yield%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_yield,aes(x=value))+
  geom_histogram(binwidth = 50)+
  labs(
    title = "Histogram for yield",
       x='Yield(Unit: gram)',
    y='Number of lands'
  )

```

Based on the histogram above, we know that the estimated plot yield for 10 individual sorghum plants is around 0 to 1000 grams where the highest possibility range is around 300 to 500 gram.\

#### 1.7 histogram for plant height

```{r}

analysis_height<-melt(newdat[,c(16)])

###clean navalue
clean_height<- analysis_height%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_height,aes(x=value))+
  geom_histogram(binwidth = 10) +
  labs(
    title = "Histogram for plant height",
       x='Average plant height(Unit: cm)',
    y='Number of lands'
  )

```

Based on the histogram above, we know range for plot height is around 30 to 300cm where usually the height will be around 70 cm.\

### 2.Explore the effect of the nitrogen treatment for leaf dimension/plant height/yield.

After exploring each variable of our dataset, the next thing we want to do is to explore the effect of the nitrogen treatment in different area like leaf dimension, plant height and yield. Hopefully, we can find some difference between different treatment.

#### 2.1 plant height

```{r}
effect_height<-newdat[,c(5,16)]
###clean navalue
clean_eff_height<- effect_height%>%
  ###cleaning na value
  drop_na(PlantHeight)

ggplot(clean_eff_height, aes(x=Treatment, y=PlantHeight))+
  geom_boxplot() +
  labs(
    title = "Plant height by nitrogen level",
       y='Average plant height(Unit: cm)',
    x='Treatment variables'
  )
ggplot(clean_eff_height, aes(x=PlantHeight, fill=Treatment))+
  geom_histogram(alpha=0.5) +
  labs(
    title = "Plant height by nitrogen level",
      x='Average plant height(Unit: cm)',
    y='Number of lands'
  )
```

Based on the boxplot above, the mean value and distribution for different treatment are all the same. Thus,we can say that there is no significant difference effect of the treatment. The histogram also support this result.\

#### 2.2 yield

```{r}
effect_yield<-newdat[,c(5,12)]
###clean navalue
clean_eff_yield<- effect_yield%>%
  ###cleaning na value
  drop_na(EstimatedPlotYield)

 
ggplot(clean_eff_yield, aes(x=Treatment, y=EstimatedPlotYield))+
  geom_boxplot() +
  labs(
    title = "Yield by treatment",
      x='Treatment variable',
    y='Estimated yield(Unit: gram)'
  )


```

Based on the boxplot above, we can see that the mean estimated yield for sufficient nitrogen level is higher than the yield for low nitrogen level which is a interesting finding. Also, we noticed that there exists some outlier points with sufficient nitrogen treatment level. The yield for these points are more than twice larger than the upper quantile yield for the low nitrogen treatment.\

#### 2.3 leaf length

Then we separate the leaf's length data and width data and plot the boxplot for different leaf by different nitrogen level to see if there exist some difference between them.

```{r}
###select required row
leaf_length<- newdat[,c(5,13,17)]
leaf_length %>% 
  rename( # here we rename the columns so things look nice in the graph later
    Flag = FlagLeafLength,
    Third = ThirdLeafLength
  ) %>% 
  pivot_longer( # then we collapse the columns for each side of the brain into a single column, with a second column holding size values
    cols = c("Flag", "Third"),
    names_to = "Side",
    values_to = "Size"
  ) %>%
  ###clean na value
  drop_na(Size)%>% # then we plot and give it a title
  ggplot(
    aes(
      x = Side,
      y = Size,
      fill = Treatment
    )
  ) + 
  geom_boxplot() +
  labs(
    title = "Box plot for flag and third leaf length by nitrogen level",
      x='Leaf type',
    y='Leaf dimension(Unit: cm)'
  )

```

Based on the box plot above, we can only say the sufficient nitrogen may have a little influence on leaf length but overall there is no difference for leaf's dimension in difference treatment since there is a large proportion of overlapping for leaf length and width between two treatment.\

#### 2.4 leaf width

```{r}
###select required row
leaf_width<- newdat[,c(5,14,18)]
leaf_width %>% 
  rename( # here we rename the columns so things look nice in the graph later
    Flag = FlagLeafWidth,
    Third = ThirdLeafWidth
  ) %>% 
  pivot_longer( # then we collapse the columns for each side of the brain into a single column, with a second column holding size values
    cols = c("Flag", "Third"),
    names_to = "Side",
    values_to = "Size"
  ) %>%
  ###clean na value
  drop_na(Size)%>% # then we plot and give it a title
  ggplot(
    aes(
      x = Side,
      y = Size,
      fill = Treatment
    )
  ) + 
  geom_boxplot() +
  labs(
 title = "Box plot for flag and third leaf width by nitrogen level",
      x='Leaf type',
    y='Leaf dimension(Unit: cm)'
  )

```

Based on the boxplot above, same as the previous analysis, we can only say the sufficient nitrogen level has no influence on flag leaf width and it may has a little negative influence on third leafwidth but overall there is no influence since there is a large overlapping between the dimension in two treatments.\

### 3. We can visualize the potential spatial effect on plant growth, indicated by plant height or flowering time, by plotting the field row and column layout and color-coding with the phenotypes.

```{r}
dat <- as.data.frame(dat)
for (i in c(1:7)) {
  dat[,i]=factor(dat[,i])
}

dat[1:5,]

ggplot(data=dat, aes(x=Row, y=Column, color=EstimatedPlotYield))+
  geom_point()+
  scale_colour_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 900,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "colour")

```

### 4. We can investigate the relationship between plant phenotypes, for example plotting 2d histograms and scatterplots bewtween each two agronomic traits.

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y, use = "complete")
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste(prefix, txt, sep="")
  if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex * abs(r))
}

dat_num <- dat[,unlist(lapply(dat, is.numeric))]  
yield_traits <- c("PaniclesPerPlot", "PanicleGrainWeight", "EstimatedPlotYield")
dat_yield <- dat[,yield_traits]
pairs(dat_yield,  upper.panel=panel.cor)
```

```{r}
library(shiny)
require(ggplot2)
require(dplyr)
require(ggpubr)
library(tidyverse)
library(ggpubr)

# Define UI for application that draws a histogram
ui <- fluidPage(
  # Application title
  titlePanel("The relationship between phenotypes and final yield"),
  
  # Sidebar with a slider input for the number of bins
  sidebarPanel(
    selectInput("phenotype",
                "Choose a phenotypic trait to investigate",
                list(`Growth` = list("DaysToBloom", "MedianLeafAngle", "FlagLeafLength", "FlagLeafWidth", 
                                     "ExtantLeafNumber","PlantHeight", "ThirdLeafLength", "ThirdLeafWidth"),
                     `Reproductive` = list("PaniclesPerPlot", "PanicleGrainWeight"))
    )
  ),
  
  # Show a plot of the generated distribution
  mainPanel(
    plotOutput("Scatterplot")
  )
)

# Define server logic required to draw a histogram
server <- function(input, output) {
  output$Scatterplot <- renderPlot({
    dat_scatter <- dat[,c(input$phenotype, "EstimatedPlotYield")]
    gg <- data.frame(Yield=dat$EstimatedPlotYield, Phenotype=dat[,input$phenotype]) %>%
      ggplot(aes(x = Phenotype, y=Yield)) + geom_point() +
      stat_smooth(method = "lm")+
      stat_regline_equation(aes(label =  paste(..rr.label.., sep = "~~~~")))+
      xlab(input$phenotype)+ylab("Yield (g)")
    print(gg)
  })
}

# Bind ui and server together
shinyApp(ui, server)
```

## Conclusion

In conclusion, when we explore each variable, we noticed that:

1.  The dimension of the third leaf is larger than the flag leaf

2.  There may exists positive relationship between flag leaf length and flag leaf width

3.  Usually it takes around 50 to 80 days for the sorghum flower bloom.The highest probability range is around 60 to 70 days.

4.  The median value of leaf angle for each 10 individual sorghum plants is around 40 to 45 degree.

5.  The grain weight for panicle is around 30 to 50 grams.

6.  The estimated plot yield for 10 individual sorghum plants is around 0 to 1000 grams where the highest possibility range is around 300 to 600 gram.

7.  The range for plot height is around 30 to 300cm where usually the height will be around 70 cm./

Then when we explore the effect of the nitrogen treatment in different area, we find out that sufficient nitrogen level has a positive influence in yield and may lead the occurrence of some extremely large yield plants. However, there is no sufficient evidence to shows that it has any relationship between plant height and leaf dimension.
