---
title: "Project presentation"
author: "Yifan Wu, Hongyu Jin"
date: today
institute: "University of Nebraska-Lincoln"
theme: UNLTheme
format: beamer
themeoptions:
  - left
  - hideothersubsections
---

## Dataset description

```{r}
library(readxl)
library(tidyverse)
dat<- read_xls("SAP2020_merged_v1.xls")
newdat<- dat[,c(-1,-3)]
```

The dataset is the agronomic traits that were measured in a sorghum field experiment with two levels of nitrogen treatment and around 400 genotypes. Hopefully we will also find the sub-population information of these genotypes as an additional categorical attribute.\

Each row represents a small pieces of land with it's unique row number and column number which contains around 10 individual sorghum plants. The explanation for each column stated below:

## List of variables

**Plot ID&Sorghum Accession:** Unique ID for each plants\

**SNPDataID：** Unique ID\

**Row&Column:** spatial data for sorghum plants\

**Block:** Blocking number\

**Treatment:** two levels of nitrogen treatment\

**DaysToBloom:** How many days until the sorghum flower bloom(unit:day)\

**MedianLeafAngle:** Median value of leaf angle for each 10 individual sorghum plants(unit:degree)\

**LeafAngleSDV:** Standard deviation of leaf angle for 10 individual sorghum plants\

**PoorStand:** If half of the sorghum plants are poor stand, if yes then the answer is Y\

**PaniclesPerPlot:** How many panicles per plot have\

**PanicleGrainWeight:** How much grain weight for each panicle(unit:gram)\

## List of variables

**EstimatedPlotYield:** The estimated plot yield for 10 individual sorghum plants(unit:gram)\

**FlagLeafLength:** The mean value of the flag leaf's length for each 10 individual sorghum plants(unit: cm)\

**FlagLeafWidth:** The mean value of the flag leaf's width for each 10 individual sorghum plants(unit: cm)\

**ExtantLeafNumber:** The number of visible leaves on the main stalk at maturity stage.\

**PlantHeight:**The mean value of the plant height for each 10 individual sorghum plants(unit: cm)\

**ThirdLeafLength:** The mean value of the third leaf's length for each 10 individual sorghum plants(unit: cm)\

**ThirdLeafWidth:** The mean value of the third leaf's width for each 10 individual sorghum plants(unit: cm)\

## Check for the properties in different variables

```{r}
library(ggplot2)
library(dplyr)
require(reshape2)
library(tidyr)


###select leaf dimention data to plot a box plot
###select required column
analysis_le<- newdat[,c(13,14,17,18)]
###clean navalue
clean_le<- analysis_le%>%
  melt()%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_le,aes(x=variable, y=value))+
  geom_boxplot()+
  labs(
    title = "Box plot for flag and third leaf dimension",
    x='Dimension variables',
    y='Leaf dimension(Unit: cm)'
  )+
  theme(text = element_text(size = 15))


```
## Correlation between leaf length and width

```{r}

Leaf_type <- c("Third leaf" = "#619CFF", "Flag leaf" = "#F8766D")
  ggplot(corr_ana)+
  geom_point(aes(x=ThirdLeafLength, y=ThirdLeafWidth, colour='Third leaf'))+
  geom_point(aes(x=FlagLeafLength, y=FlagLeafWidth, colour='Flag leaf'))+
  labs(
    title = "Scatter plot for flag and third leaf's dimension",
    x='leaf Length (Unit: cm)',
    y='leaf width (Unit: cm)'
  )+
    scale_color_manual(values = Leaf_type)+
  theme(text = element_text(size = 15))
```

## Check for the properties in different variables
```{r}
##create some histogram for each variable to have a basic understand for each variable

###blooming date
analysis_bloom<-melt(newdat[,c(6)])

###clean navalue
clean_bloom<- analysis_bloom%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_bloom,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "Histogram for blooming day",
       x='Blooming time(Unit: day)',
    y='Counts of lands'
  )+
  theme(text = element_text(size = 15))


```

## Check for the properties in different variables
```{r}

analysis_angle<-melt(newdat[,c(7)])

###clean navalue
clean_angle<- analysis_angle%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_angle,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "Histogram for median leaf angle",
       x='Leaf angle(Unit:degree)',
    y='Count of lands'
  )+
  theme(text = element_text(size = 15))

```

## Check for the properties in different variables
```{r}

analysis_panicle<-melt(newdat[,c(10)])

###clean navalue
clean_panicle<- analysis_panicle%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_panicle,aes(x=value))+
  geom_histogram(binwidth = 2)+
  labs(
    title = "Histogram for panicle amount",
       x='Panicle amount',
    y='Number of lands'
  )+
  theme(text = element_text(size = 15))

```
## Check for the properties in different variables

```{r}

analysis_grain<-melt(newdat[,c(11)])

###clean navalue
clean_grain<- analysis_grain%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_grain,aes(x=value))+
  geom_histogram(binwidth = 5)+
  labs(
    title = "Histogram for grain weight",
       x='Average grain Weight(Unit: gram)',
    y='Number of lands'
  )+
  theme(text = element_text(size = 15))

```

## Check for the properties in different variables
```{r}

analysis_yield<-melt(newdat[,c(12)])

###clean navalue
clean_yield<- analysis_yield%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_yield,aes(x=value))+
  geom_histogram(binwidth = 50)+
  labs(
    title = "Histogram for yield",
       x='Yield(Unit: gram)',
    y='Number of lands'
  )+
  theme(text = element_text(size = 15))

```
## Check for the properties in different variables

```{r}

analysis_height<-melt(newdat[,c(16)])

###clean navalue
clean_height<- analysis_height%>%
  ###cleaning na value
  drop_na(value)

###plot box plot
ggplot(clean_height,aes(x=value))+
  geom_histogram(binwidth = 10) +
  labs(
    title = "Histogram for plant height",
       x='Average plant height(Unit: cm)',
    y='Number of lands'
  )+
  theme(text = element_text(size = 15))

```

## Explore the effect of the nitrogen treatment with plant height

```{r}
effect_height<-newdat[,c(5,16)]
###clean navalue
clean_eff_height<- effect_height%>%
  ###cleaning na value
  drop_na(PlantHeight)

ggplot(clean_eff_height, aes(x=Treatment, y=PlantHeight))+
  geom_boxplot() +
  labs(
    title = "Plant height by nitrogen level",
       y='Average plant height(Unit: cm)',
    x='Treatment variables'
  )+
  theme(text = element_text(size = 15))

```

## Explore the effect of the nitrogen treatment with yield

```{r}
effect_yield<-newdat[,c(5,12)]
###clean navalue
clean_eff_yield<- effect_yield%>%
  ###cleaning na value
  drop_na(EstimatedPlotYield)

 
ggplot(clean_eff_yield, aes(x=Treatment, y=EstimatedPlotYield))+
  geom_boxplot() +
  labs(
    title = "Yield by treatment",
      x='Treatment variable',
    y='Estimated yield(Unit: gram)'
  )+
  theme(text = element_text(size = 15))


```

## Explore the effect of the nitrogen treatment with leaf length

```{r}
###select required row
leaf_length<- newdat[,c(5,13,17)]
leaf_length %>% 
  rename( # here we rename the columns so things look nice in the graph later
    Flag = FlagLeafLength,
    Third = ThirdLeafLength
  ) %>% 
  pivot_longer( # then we collapse the columns for each side of the brain into a single column, with a second column holding size values
    cols = c("Flag", "Third"),
    names_to = "Side",
    values_to = "Size"
  ) %>%
  ###clean na value
  drop_na(Size)%>% # then we plot and give it a title
  ggplot(
    aes(
      x = Side,
      y = Size,
      fill = Treatment
    )
  ) + 
  geom_boxplot() +
  labs(
    title = "Box plot for flag and third leaf length by nitrogen level",
      x='Leaf type',
    y='Leaf dimension(Unit: cm)'
  )+
  theme(text = element_text(size = 15))

```

## Explore the effect of the nitrogen treatment with leaf width
```{r}
###select required row
leaf_width<- newdat[,c(5,14,18)]
leaf_width %>% 
  rename( # here we rename the columns so things look nice in the graph later
    Flag = FlagLeafWidth,
    Third = ThirdLeafWidth
  ) %>% 
  pivot_longer( # then we collapse the columns for each side of the brain into a single column, with a second column holding size values
    cols = c("Flag", "Third"),
    names_to = "Side",
    values_to = "Size"
  ) %>%
  ###clean na value
  drop_na(Size)%>% # then we plot and give it a title
  ggplot(
    aes(
      x = Side,
      y = Size,
      fill = Treatment
    )
  ) + 
  geom_boxplot() +
  labs(
 title = "Box plot for flag and third leaf width by nitrogen level",
      x='Leaf type',
    y='Leaf dimension(Unit: cm)'
  )+
  theme(text = element_text(size = 15))

```
